---
title: "alpscarf"
author: "Chia-Kai Yang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{alpscarf tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Initial Setup
Import the library we suggest to use and the package _alpscarf_.
```{r setup, echo = FALSE}
library(tidyverse)
import::from(cowplot, plot_grid)
import::from(magrittr, "%<>%")
library(alpscarf)
```

## Data Import
Import example data. The dataset _reduced_dwell.RData_ includes __AOI visit sequences__ for 17 participants (P1 - P8, P10 - P18). The dataset _aoi_sequence.RData_ contains the __expected visit order__ of each AOI, as well as the __color association__ of AOIs.
```{r get dataset}
# specify dataset name & path
dataset_path = "../data"
example_dataset <- file.path(dataset_path, "reduced_dwell.RData")
expected_aoi_sequence <- file.path(dataset_path, "aoi_sequence.RData")

# load data  
load(example_dataset)
load(expected_aoi_sequence)

# put AOIs in sorted order
aoi_names_pages_seq %<>%
  arrange(AOI_order)
```

The __AOI visit sequences__ contains at least 3 columns: "AOI" "dwell_duration" "p_name" 
```{r}
head(lsa_reduced_dwell_df)
```


The __expected visit order__ contains at least two columns: "AOI" and "AOI_order".
```{r}
head(aoi_names_pages_seq)
```
In above example dataset, we merge the __expected visit order__ with and __color association__, resulting the an extra column ```color```.

## Generate Alpscarf
When the data is ready, users can generate Alpscarf in following steps:

1. Define colors for AOIs

The association between AOIs and colors MUST be an 1-to-1 mapping; i.e., n AOIs require n colors. An exmaple of such mapping is included in the dataset _aoi_sequence.RData_.
```{r color association}
# define colors
my_palette <- 
  aoi_names_pages_seq$color[]
```

2. Translate visit sequence into bar height/width 

Two funcitons are involved in this step:

  * _alpscarf_height_trans_: the function calculates the confomity/revisiting scores and translate them into bar height
  * _alpscarf_width_trans_: the function calculates the position of bars where bar width equals dwell duration (in logarithmic scale)
```{r spec translation}
# generate Alpscarf dataset
lsa_reduced_dwell_alp_df <- 
  lsa_reduced_dwell_df %>%
  alpscarf_height_trans(., aoi_names_pages_seq) %>%
  alpscarf_width_trans()
```
Note that some parameters are specifiable in the function _alpscarf_height_trans_. Users can choose from linear mode (```LINEAR_MODE = TRUE```) and exponential mode (```LINEAR_MODE = FALSE```), which controls the contrast of mountain height (i.e., make high mountains even higher, and vice versa). In exponential mode, users can also specify the base (```base_factor```) and the scale (```scale_factor```) of exponent.

3. Generate Alpscarf plot

Now the dataset is ready for generating Alpscarf. In this example, we generate one Alpscarf per participant.
```{r alpscarf plot}
# initialise the list storing Alpscarfs, one plot per participant
lsa_scarf_vis <- list()
# specify plot height
plot_height <- max(lsa_reduced_dwell_alp_df$seq_bar_length)

for(a_p_name in unique(lsa_reduced_dwell_alp_df$p_name)){
  df_p <-
    lsa_reduced_dwell_alp_df %>%
    filter(p_name == a_p_name) 
  
  # for indexing
  a_p_nr <- 
    strsplit(a_p_name,"P")[[1]][2] %>%
    as.numeric()
  
  # specify the color association of AOIs
  aoi_name_in_order <- 
    aoi_names_pages_seq$AOI[]
  df_p$AOI <- factor(df_p$AOI, levels = aoi_name_in_order)
  
  # Alpscarf plot generation
  lsa_scarf_vis[[a_p_nr]] <- alpscarf_plot_gen(df_p, my_palette, TRANSITION_FOCUS = TRUE, ALPSCARF_EN = TRUE,  max_y_value = plot_height, TITLE_EN = FALSE)
}

# plot Alpscarf for all participants
plot_grid(lsa_scarf_vis[[7]], lsa_scarf_vis[[14]], lsa_scarf_vis[[11]], lsa_scarf_vis[[17]], lsa_scarf_vis[[15]], lsa_scarf_vis[[5]], lsa_scarf_vis[[1]], lsa_scarf_vis[[13]], lsa_scarf_vis[[16]], lsa_scarf_vis[[6]], lsa_scarf_vis[[4]], lsa_scarf_vis[[10]], lsa_scarf_vis[[12]], lsa_scarf_vis[[2]], lsa_scarf_vis[[3]], lsa_scarf_vis[[18]], lsa_scarf_vis[[8]], align = "hv", ncol = 2)
# use ggsave to save plots
ggsave(file = "temp.png", plot = last_plot(), width = 49, height = 49, device = "png", bg = "transparent")
```
Note that, in above example, we generate Alpscarf in __transition-focus__ mode. Users can also generate Alpscarf in __duration-focus__ mode (```TRANSITION_FOCUS = FALSE```), or even traditional scarf plot (```ALPSCARF_EN = FALSE```).

Besides, ```max_y_value``` specify the height of each plot. To make all Alpscarf comparable, generating Alpscarfs in the same height is highly recommended.


4. (bonus) Distance measures of AOI visit sequences

__This part is NOT necessary to generate Alpscarf plot__

Additionally, users can get a few __descriptive stats__ (e.g., order_following, rereading ratio), and __edit distance__ between each AOI sequence and the expected visit order. 
```{r edit distance}
# generate stats of Alpscarf; NOT required in generating Alpscarf.
stat_table_df <- alpscarf_calculate_statistics(lsa_reduced_dwell_alp_df, aoi_names_pages_seq)

head(stat_table_df)
```
